---
title: "Seattle Pets"
author: "Nancy Huynh"
date: '2019-03-26'
output: 
  github_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Seattle Pet Names
Quick look at pets registered in the Seattle. The data originates from [Seattle's Open Data](https://data.seattle.gov/Community/Seattle-Pet-Licenses/jguv-t9rb), and was also made available in the #TidyTuesday [repo](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-03-26). 

##Libraries & Data Import
```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(skimr)

seattle_pets <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-03-26/seattle_pets.csv")

glimpse(seattle_pets)

#converting date string to Date type after viewing the glimpse
seattle_pets <- seattle_pets %>%
  mutate(license_issue_date = mdy(license_issue_date))
```

```{r echo=FALSE, include=FALSE}
## my theme object from previous TidyTuesday
nh_theme <- theme(panel.background = element_rect(fill = "white"),
          strip.background = element_blank(),
          strip.text = element_text(hjust = 0, size = 14, color = "grey25", face = "bold"),
          panel.spacing = unit(2, "lines"),
          axis.line = element_line(color = "grey85"),
          panel.grid.major.y = element_line(color = "grey75", linetype = "dotted"),
          legend.key = element_blank(),
          legend.key.height = unit(1, "cm"),
          legend.title = element_text(size = "10"),
          plot.title = element_text(size = 15, color = "grey15", face = "bold"),
          plot.subtitle = element_text(size = 12, color = "grey30"),
          text = element_text(color = "grey40"))
```

###Quick Look at the Data
There's quite a bit of data available, and it goes pretty far back, but it doesn't look like tracking officially started until about 2014. And it's not just cats and dogs either. Some people have registered their pigs and goats as pets.
```{r}
seattle_pets %>%
  arrange(license_issue_date) %>%
  head(12)

skim(seattle_pets)

seattle_pets %>%
  count(species)

```


###Cat or Dog Names
Curious to know if there's some names that are more "dog" and some that are more "cat". Interestingly Luna is equally popular with cats and dogs. I would have guessed its a more "cat" name 'cause of [Sailor Moon's cat](https://yattatachi.com/top-6-iconic-and-adorable-pets-in-anime-history). Also no surprise that there are some names that people only give to cats, i.e. Kitty. Although someone ought to break the mould with that one...
```{r catordogname, fig.width=10}
seattle_pets %>% 
  filter(species %in% c("Cat", "Dog"),
         license_issue_date > "2017-12-31",
         !is.na(animals_name)) %>%
  group_by(animals_name, species) %>%
  summarise(n = n()) %>%
  group_by(species) %>%
  mutate(prop = n / sum(n)) %>%
  filter(n > 10) %>%
  select(animals_name, species, prop) %>%
  spread(key = species, value = prop, fill = 0) %>%
  ggplot(aes(x = Dog, y = Cat)) +
  geom_text(aes(label = animals_name, size = Cat + Dog), show.legend = FALSE) +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(labels = scales::percent) +
  nh_theme +
  labs(x = "% of Dogs",
       y = "% of Cats",
       title = "Percent of Cats and Dogs with Specific Names",
       subtitle = "Includes names with over 10 registrations in 2018",
       caption = "Data: Seattle Open Data; Graphic: @nh_writes") +
  geom_abline(slope = 1, intercept = 0, linetype = 2)

```
*Pretty messy chart here!*

###First Letter of Pet Names
Roughly from the chart above we can see that names that start with "L" are quite popular. Here's a closer look. "M" tops the list for both cats and dogs. "L" is definitely in the top 5. The top five letters are the same for both cats in dogs, just in slightly different order.
```{r}
seattle_pets %>%
  mutate(first_letter = str_to_upper(substr(animals_name, 1, 1))) %>%
  group_by(first_letter, species) %>%
  filter(first_letter %in% LETTERS,
         species %in% c("Cat", "Dog"),
         license_issue_date > "2017-12-31") %>%
  summarize(n = n()) %>%
  group_by(species) %>%
  mutate(prop = (n / sum(n))*100, rank = rank(-prop)) %>%
  filter(rank <= 5) %>%
  ungroup() %>%
  arrange(species, desc(prop)) %>%
  transmute(`Pet Type` = species,
            `Rank` = rank,
            `Name Starts With` = first_letter,
            `% With Name` = prop) %>%
  head(10) %>% knitr::kable()
```

Here's a chart with all the letters of the alphabet.
```{r alphapets, fig.width=10, fig.height=10}

seattle_pets %>%
  mutate(first_letter = str_to_upper(substr(animals_name, 1, 1))) %>%
  group_by(first_letter, species) %>%
  filter(first_letter %in% LETTERS,
         species %in% c("Cat", "Dog"),
         license_issue_date > "2017-12-31") %>%
  summarize(n = n()) %>%
  group_by(species) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = fct_rev(first_letter), y = prop, fill = factor(species))) +
  geom_col(position = "dodge", width = 0.5) + 
  scale_x_discrete(name = "") +
  scale_y_continuous(labels = scales::percent, name = "Proportion of Pet Type with Names", expand = c(0,0), limits = c(0, 0.15), breaks = seq(from = 0, to = 0.15, by = 0.025)) +
  scale_fill_manual(name = "", breaks = c("Dog", "Cat"), values = c("Cat" = "violet","Dog" = "dodgerBlue")) +
  coord_flip() +
  nh_theme +
  theme(panel.grid.major.x = element_line(colour = "grey80", linetype = 1, size = 0.25),
        panel.grid.major.y = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "right") +
  labs(title = "First Letter of Pet Names",
       subtitle = "2018 license registrations in Seattle",
       caption = "Data: Seattle Open Data; Graphic: @nh_writes")
```

##After Thoughts
Learned a few more R tricks with this one, including how to make a quick `kable()` table using `transmute()` to clean up the column names. Also tried out `substr()` for the first time. I wonder how good a decision tree or random forest model would be using letter of first name and length of name and year of registration would be at predicting the type of animal (cat vs dog)... it would be a pretty silly model to build anyway, lol.