---
title: "Tennis"
author: "Nancy Huynh"
date: '2019-04-11'
output: 
  github_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tennis Grand Slams
This week's TidyTuesday dataset looks at Grand Slam Tournaments since 1968. I used to follow tennis more closely but haven't kept up with the players and the games in recent years. This should be fun! Dataset can be found in the [TidyTuesday repo](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-04-09).

## Libraries and Data Import

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

```{r echo=FALSE, include=FALSE}
## my theme object from previous TidyTuesday
nh_theme <- theme(panel.background = element_rect(fill = "white"),
          strip.background = element_blank(),
          strip.text = element_text(hjust = 0, size = 12, color = "grey25", face = "bold"),
          panel.spacing = unit(2, "lines"),
          axis.line = element_line(color = "grey85"),
          panel.grid.major.y = element_line(color = "grey75", linetype = "dotted"),
          legend.key = element_blank(),
          legend.key.height = unit(1, "cm"),
          legend.title = element_text(size = "10"),
          plot.title = element_text(size = 15, color = "grey15", face = "bold"),
          plot.subtitle = element_text(size = 12, color = "grey30"),
          text = element_text(color = "grey40"))
```

I filtered out 2019 since the season has just started.
```{r message=FALSE, warning=FALSE}
grand_slams <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-04-09/grand_slams.csv") %>%
  filter(year < 2019)

player_dob <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-04-09/player_dob.csv")
```

## How many Grand Slams did players win at each age of their life?

Calculate the number of grand slams won at each player's age. Note that date of the Grand Slam matches are estimates.

```{r}
age_slams_comb <- left_join(grand_slams, player_dob, by = c("name")) %>% 
  mutate(age = floor(as.numeric(tournament_date - date_of_birth)/365.2425)) %>%
  group_by(name, age, gender) %>% 
  summarize(counts = n()) %>% 
  group_by(name) %>% 
  mutate(total_wins = cumsum(counts)) %>% 
  arrange(desc(total_wins))
```

Pull out the top 10 players based on total life-time Grand Slam wins for each gender.
```{r}
top_players <- age_slams_comb %>%
  group_by(name, gender) %>%
  summarise(max_total_wins = max(total_wins)) %>%
  ungroup() %>%
  group_by(gender) %>%
  top_n(10, max_total_wins) %>%
  ungroup()%>%
  select(name) %>% 
  pull()

plot_top_players <- age_slams_comb %>%
  filter(name %in% top_players)

```

Plot the number of Grand Slam wins at each age for the top players. Margaret Court is the only player to win 4 Grand Slams in one year of her life (at age 27). Although this should be taken with a grain of salt since the match dates are estimates in this data set. Meanwhile, Serena Williams and Roger Federer are still rocking it well into their late 30s.

```{r age_wins, fig.width=10, fig.height=9}

ggplot(plot_top_players, aes(x = fct_reorder(name, counts, sum), y = age, fill = counts)) +
  geom_tile() +
  coord_flip() +
  facet_wrap(~gender, nrow = 2, scales = "free_y") +
  scale_fill_continuous(low = "yellow", high = "green4", name = "Grand Slams Won") +
  nh_theme +
  theme(panel.grid.major.x = element_blank(),
        legend.position = "top", legend.direction = "horizontal") +
  labs(x = "",
       y = "Age",
       title = "Margaret Court is the only player to win 4 Grand Slams in one year of her life*",
       subtitle = "Grand Slams (1968-2018) won by age; sorted by total Grand Slams won",
       caption = "*Note: Match dates are estimates; so take this with a grain of salt \n 
       Data: Wikipedia | Graphic: @nh_writes")

#ggsave("tennis_grand_slams.png", device = "png", units = "cm", width = 29, height = 21, dpi = "retina")

```

