---
title: "Women in the Workforce"
author: "Nancy Huynh"
date: '2019-03-06'
output: 
  github_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Women in the Workforce (US)
This week's TidyTuesday dataset is aptly in line with International Women's Day (March 8) and Women's History month. Find the data here on the [tidytuesday repo](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-03-05). There's three data files, the one I'm using below comes from the US Census Bureau. *Side note: This might be an apt time to highlight that Viola Desmond, Canadian activist, is [featured on our $10 bill](https://www.theglobeandmail.com/canada/article-viola-desmond-10-bill-explainer/).* `r emo::flag("Canada")`

## Libraries and Data Import
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(skimr)
library(scales)
library(RColorBrewer)
```

```{r warning=FALSE}
jobs_gender <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-03-05/jobs_gender.csv")
```


### Quick Look at Data and Variables
Recently learned about the `skimr` package at the kickoff R Ladies Toronto meetup this week from [Sharla Gelfand](https://sharla.party/talks/rladies-rug-kickoff.html#1). Note: the `skimr` package is maintained by by Elin Waring and Michael Quinn. Decided to give `skimr` a try. It was handy to see the number of unique values for the `major_category`, `minor_category`, and `occupation` variables. Also handy is the missing values column.
```{r}
glimpse(jobs_gender)
skim(jobs_gender)

# take a look at which occupations had missing values for the total_earnings_female and total_earnings_male variables
jobs_gender %>%
  filter(is.na(total_earnings_female)) %>%
  group_by(minor_category) %>%
  summarise(n = n())

jobs_gender %>%
  filter(is.na(total_earnings_male))

```

The `skimr` output showed 522 unique so I'm not going to print that out, but let's take a look at the major and minor categories. 
```{r}
# unique major_category
unique(jobs_gender$major_category)

# unique minor_categories
unique(jobs_gender$minor_category)



```

### Where the "Analysts" at?
I've been various types of "analysts" over the years so I'm curious to see what the earnings are like for occupations with "analyst" in the occupation name. Since there's no existing category for Analysts, I just filtered them out using `grepl()`. There are 9 unique occupations with Analyst in the name. Interestingly the vague occupation of "Business Analyst" isn't one of them. I did a quick search to see how the US Census came up with the list of occupations but didn't find anything. My best guess is that work titles like "Business Analyst" are lumped into one of the listed occupations—probably "Management analysts". In reality there are probably hundreds of titles with "Analyst", search LinkedIn if you don't believe me.
```{r}
# find all occupations with "analyst"" in the name
analyst_jobs <- jobs_gender %>%
  filter(grepl("analyst", occupation))  %>%
  select(year, occupation, minor_category, total_earnings_female, total_earnings_male, total_workers)

analyst_jobs %>%
  arrange(desc(total_workers))

unique(analyst_jobs$occupation)

```

Let's see how the median earnings of these "Analyst" occupations have changed over time. And split by female vs male. There's some trends here. But the one that this particular, slightly messy, chart highlights just because of the position of the lines is the difference for earnings between female and male for 'Market research analysts and marketing specialists'.
```{r fig.width=10}

nh_theme <- theme(panel.background = element_rect(fill = "white"),
          strip.background = element_blank(),
          strip.text = element_text(hjust = 0, size = 14, color = "grey25", face = "bold"),
          panel.spacing = unit(2, "lines"),
          axis.line = element_line(color = "grey85"),
          legend.key = element_blank(),
          legend.key.height = unit(1, "cm"),
          legend.title = element_text(size = "10"),
          plot.title = element_text(size = 15, color = "grey15", face = "bold"),
          plot.subtitle = element_text(size = 12, color = "grey30"),
          text = element_text(color = "grey40"))

analyst_jobs %>%
  select(year, occupation, total_earnings_female, total_earnings_male) %>%
  gather(key = "gender", value = "total_earnings", -year, -occupation) %>%
  mutate(
    gender = case_when(
      gender == "total_earnings_female" ~ "Female",
      gender == "total_earnings_male" ~ "Male"
    )) %>%
  ggplot(aes(x = year, y = total_earnings, color = fct_reorder(str_wrap(occupation, 20), -total_earnings))) +
    geom_line() +
    # show both x and y axis when faceting by setting to scales = "free"
    facet_wrap(~ gender, scales = "free") +
    scale_color_brewer(palette = "Paired", name = "Occupations") +
    # set limits to y-axis so that both y axis line up with each other
    # format axis labels to show $
    scale_y_continuous(labels = scales::comma_format(prefix = "$"), limits = c(40000, 101000), name = "Median Earnings") +
    # give some padding around the year axis, but also make sure the breaks are on the years
    scale_x_continuous(breaks = c(2013:2016), expand = c(0, 0.5), name = "Year") +
    nh_theme +
    labs(title = "Earnings for Different Types of Analysts by Gender (2013-2016)",
       subtitle = "All occupations with \"analyst\" in the title",
       caption = "Data source: US Census Bureau // #TidyTuesday // @nh_writes")
```

It's hard to tell exactly what the earnings gap is between female and male for each of the positions so here's a dumbbell chart of 2016 earnings for the 9 analyst occupations. I sorted it by median earnings gap between female and male. It was hard to tell in the line chart what the gap is, except for those 'Market research analysts and marketing specialists'. In this one we can see that in 2016 the biggest gap is between female and male financial analysts. For the most part, the gap is larger for occupations with earnings toward the higher end. *Note: Dumbbell chart inspired by tweet from [\@rcammisola](https://twitter.com/rcammisola/status/1103351691993382912)*

```{r fig.width=10}
library(ggalt) # for geom_dumbbell

colors <- brewer.pal(4, "Paired")

analyst_jobs %>%
  filter(year == 2016) %>%
  mutate(occupation = fct_reorder(str_wrap(occupation, 20), (total_earnings_male - total_earnings_female))) %>%
  ggplot(aes(x = total_earnings_female, xend = total_earnings_male, y = occupation, yend = occupation)) +
  geom_dumbbell(colour_x = colors[4], colour_xend = colors[2], size_x = 3, size_xend = 3, dot_guide = TRUE, dot_guide_size = 0.25) +
  geom_text(data = filter(analyst_jobs, year == 2016, occupation == "Financial analysts"), aes(x = total_earnings_female, y = occupation, label="Female"), vjust = -1, color = colors[4], fontface = "bold") +
  geom_text(data = filter(analyst_jobs, year == 2016, occupation == "Financial analysts"), aes(x = total_earnings_male, y = occupation, label="Male"), vjust = -1,  color = colors[2], fontface = "bold") +
  scale_x_continuous(labels = scales::comma_format(prefix = "$"), limits = c(45000, 101000), breaks = seq(50000, 100000, 10000), name = "Median Earnings in 2016") +
  scale_y_discrete(name = "Occupations") +
  nh_theme +
  theme(panel.grid.major.x = element_line(size = 0.1, color = "grey85")) +
  labs(title = "Earnings for Different Types of Analysts by Gender (2016)",
      subtitle = "Sorted Descending by Median Earnings Gap",
      caption = "Data source: US Census Bureau // #TidyTuesday // @nh_writes")

```

## BTW and Thanks `r emo::ji("pray")`
I had no idea what a dumbbell chart was until I started checking out what other people were doing for [#TidyTuesday](https://twitter.com/search?q=%23TidyTuesday&src=tyah). Also found out what a slope graph is, maybe I'll try that out for another TidyTuesday. Big thanks to the #TidyTuesday and #rstats community!