---
title: "Ramen Ratings"
author: "Nancy Huynh"
date: '2019-06-04'
output: 
  github_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Ramen Ratings from The Ramen Rater
This week I take a stab at LASSO regression using the week 23  [#TidyTuesday](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-06-04) data. The original data comes from [The Ramen Rater](https://www.theramenrater.com/resources-2/the-list/), which is indeed ratings for ramen (the "instant noodle" type). I first saw LASSO regression from [Dave Robinson's YouTube screencast](https://www.youtube.com/watch?v=qirKGdQvy9U) of his analysis on a TidyTuesday dataset earlier this year. Other resources I used are below.

## Data Import and Libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(glmnet)
library(broom)

ramen_ratings <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-04/ramen_ratings.csv", col_types = "ifcffd")

summary(ramen_ratings)

```

```{r echo=FALSE, include=FALSE}
## my theme object from previous TidyTuesday
nh_theme <- theme(panel.background = element_rect(fill = "white"),
          strip.background = element_blank(),
          strip.text = element_text(hjust = 0, size = 12, color = "grey25", face = "bold"),
          panel.spacing = unit(2, "lines"),
          axis.line = element_line(color = "grey85"),
          panel.grid.major.y = element_line(color = "grey75", linetype = "dotted"),
          legend.key = element_blank(),
          legend.key.height = unit(1, "cm"),
          legend.title = element_text(size = "10"),
          plot.title = element_text(size = 15, color = "grey15", face = "bold"),
          plot.subtitle = element_text(size = 12, color = "grey30"),
          text = element_text(color = "grey40"))
```

## What are the ratings across brand, style, and country of origin?
Since there are numerous countries and brands I use `fct_lump` to grab the top 10 and lump the remaining into other.
```{r ramen_boxplots}
ramen_boxplots <- function(ramen_df, ramen_col) {
  ramen_col <- enquo(ramen_col)
  
  ramen_df %>%
    ggplot(aes(x = fct_relevel(fct_reorder(fct_lump(!!ramen_col, n = 10), stars, na.rm = TRUE), "Other", after = 0), y = stars)) +
    geom_boxplot(na.rm = TRUE) + 
    coord_flip() +
    labs(y = "Star Rating",
         x = str_to_title(rlang::as_label(ramen_col)))
}

ramen_boxplots(ramen_ratings, brand)
ramen_boxplots(ramen_ratings, country)
ramen_boxplots(ramen_ratings, style)
```

## What factors influences the star rating of the ramen?
Going to try a lasso regression without engineering any extra features. But, some ideas for engineered features from the `variety` variable could be spicy (i.e. words like "Spicy", "Chili") or noodle type (i.e. words like "udon"", "vermicelli").
```{r ramen_lasso}

#training data set
ramen_train  <- ramen_ratings %>%
  sample_frac(0.80) %>%
  na.omit()

#testing data set
ramen_test <- ramen_ratings %>%
  anti_join(ramen_train, by = "review_number") %>%
  na.omit()

#create matrix of predictors
features <- model.matrix(stars ~ brand + style + country, ramen_train)[, -1]

#determine which value of lambda to use
cv.lasso <- cv.glmnet(x = features, y = ramen_train$stars, alpha = 1)
plot(cv.lasso)

cv.lasso$lambda.min
cv.lasso$lambda.1se

#choosing lambda.1se as this seems to be the typical choice so as to not overfit
ramen_coefs <- cv.lasso$glmnet.fit %>%
  tidy() %>%
  filter(lambda == cv.lasso$lambda.1se) %>%
  arrange(desc(estimate))
```

```{r ramen_coefs}
ramen_coefs %>%
  filter(term != "(Intercept)") %>%
  top_n(30, abs(estimate)) %>%
  mutate(term = str_replace(term, "brand", "brand: "),
         term = str_replace(term, "country", "country: "),
         term = str_replace(term, "style", "style: ")) %>%
  ggplot(aes(x = fct_reorder(term, -estimate), y = estimate)) +
  geom_col() +
  coord_flip() +
  labs(y = "Coefficient",
       x = "",
       title = "Largest coefficients to predict ramen ratings",
       subtitle = "Based on LASSO regression",
       caption = "Data: The Ramen Rater \n @nh_writes") +
  nh_theme

#ggsave("ramen_coefficients.png", device = "png", units = "cm", width = 29, height = 21, dpi = "retina")
```

## Things to do later
This is by no means a complete analysis; I'm just trying out LASSO regression using this fairly simple dataset. After looking at the top 30 coefficients I should somehow limit the brands to those that occur more than say 5 times. For example the largest coefficient is brand "Roland", which only occurs 2 times in the whole dataset. I should also see how well the model fairs against the test data set I split out called `ramen_test`.

## References
* [Lasso Regression: Simple Definition](https://www.statisticshowto.datasciencecentral.com/lasso-regression/)
* [Penalized Logistic Regression Essentials in R: Ridge, Lasso and Elastic Net](http://www.sthda.com/english/articles/36-classification-methods-essentials/149-penalized-logistic-regression-essentials-in-r-ridge-lasso-and-elastic-net/)
