---
title: "France Trains"
author: "Nancy Huynh"
date: '2019-02-28'
output: 
  github_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# France Trains Data - TidyTuesday
With this TidyTuesday dataset I learned how to make a circular network chart. I also learned some of the possible downsides to this type of chart particularly when using colored arcs to represent a variable and the issue of overlapping arcs. For the dataset head over to [TidyTuesday](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-02-26)

## Data Imports and Libraries
```{r, warning=FALSE, message=FALSE}
#LOAD LIBRARIES
library(tidyverse)
library(igraph) # for circular network graph
library(ggraph) # for circular network graph
```

```{r, message=FALSE}
#LOAD DATA
trains <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-02-26/small_trains.csv")

```

### Quick Look at Data
Check out the structure of the data and what the most recent data is. Latest data is from November 2018.
```{r}
str(trains, give.attr = FALSE)

trains %>% 
  arrange(desc(year), desc(month)) %>%
  head()
```

### Looking at Total Number of Trips in 2018
Taking a look at total number of trips variable `total_num_trips` in 2018. The route with the most trips in 2018 is from BORDEAUX ST JEAN to PARIS MONTPARNASSE. Note from the printout for the tibble size that there were 130 routes in 2018.
```{r}
routes_2018 <- trains %>% filter(year == 2018) %>%
  group_by(departure_station, arrival_station) %>%
  summarise(total_trips = sum(total_num_trips))

routes_2018 %>%  
  arrange(total_trips)
  
```

I saw a pretty cool circular network chart on this dataset from [\@WireMonkey](https://twitter.com/WireMonkey/status/1100932771122421762) and wanted to learn how to make something similar just for the total trips in 2018 from departure to arrival stations. After figuring out how to do it something bothered me with the chart.  My issue with this chart is that it's hard to distinguish the correct color for overlapping arcs. For example Paris Nord to Lille, and Lille to Paris Nord, the arcs overlap and I can't tell if the color of the arc represents the former or latter departure-arrival pair. It loooks pretty cool, and we can tell quickly which stations have the most routes, but not sure if its the best to determine total trips.

```{r fig.width=10, fig.height=10}
routes_2018 <- routes_2018 %>%
  arrange(departure_station)

## Function to create circular network graph so I can reuse it below instead of copying pasting everything
circle_graph <- function(data) {
net_graph <- graph_from_data_frame(data, directed = TRUE)

## Figure out the angles for each label so they are easily legible
# create layout to access index to determine angles
layout <- create_layout(net_graph, layout = "linear", circular = TRUE)
# calculate angles and adjustments for labels
V(net_graph)$angle <- 90 - 360 * (layout$ggraph.index-0.5) / max(layout$ggraph.index)
V(net_graph)$hjust <- ifelse(V(net_graph)$angle < -90, 1.1, -0.1)
V(net_graph)$angle <- ifelse(V(net_graph)$angle < -90, V(net_graph)$angle+180, V(net_graph)$angle)

## Create circular network graph
chart <- net_graph %>%
  ggraph(layout = "linear", circular = TRUE) +
  geom_node_point(alpha = 0) +
  geom_node_text(aes(label = str_wrap(name, 8), angle = angle, hjust = hjust), size = 2, alpha = 0.5) +
  scale_edge_color_gradient(low = "grey85", high = "navyBlue") +
  coord_fixed() +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.background = element_rect(fill = "white"),
        legend.text = element_text(color = "grey50"),
        legend.title = element_text(color = "grey50"),
        legend.position = "left")

return(chart)
}

circle_graph(routes_2018) +
  geom_edge_arc(aes(color = total_trips), arrow = arrow(length = unit(0.10, "inches"))) +
  labs(edge_color = "Total Trips",
       title = "Total Trips from Departure to Arrival Stations (2018)",
       subtitle = "All Stations Included")
```

The only reason that I'm "ok" with the chart is that the largest absolute difference between a departure-arrival pair and its reverse pair is 5508. This might be a fairly large number, but when it comes to the color scaling a difference of 5000 trips is not *that* distinguishable. 
```{r}
## Find the reverse number of trips and calculate the absoluate difference
routes_2018_dup <- routes_2018 %>%
  rename(rev_trips = total_trips)

left_join(routes_2018, routes_2018_dup, by = c("arrival_station" = "departure_station", "departure_station" = "arrival_station")) %>%
  mutate(difference = abs(total_trips - rev_trips)) %>%
  arrange(desc(difference))

```

The alternative might be to calculate the total trips going in and out of the two stations and then the colored arcs would represents all trips coming and going between the two stations. Although apart from the change in basically doubling the values on the color scale the charts look the same.
```{r fig.width=10, fig.height=10}

routes_2018_new <- routes_2018 %>%
  left_join(routes_2018_dup, by = c("arrival_station" = "departure_station", "departure_station" = "arrival_station")) %>%
  mutate(total_trips = total_trips + rev_trips) %>%
  select(departure_station, arrival_station, total_trips) %>%
  arrange(departure_station)

circle_graph(routes_2018_new) +
    geom_edge_arc(aes(color = total_trips)) +
  labs(edge_color = "Total Trips",
         title = "Number of Trips In and Out of Stations (2018)",
         subtitle = "Arc color represents sum of trips to and from station nodes")
```

